
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Setup Dev Environment w/ defaults
```{r}
# Set working directory
setwd("C:/Projects/Coursera/RR")
# Install required packages (if needed, may require prompt answer)
#install.packages("data.table")

# Load required libraries, ggplot to be used for all plots.
library(knitr)
library(data.table)
library(ggplot2)

# Set default chunk code
opts_chunk$set(echo = TRUE, results = 'hold')

```



## Loading and preprocessing the data
Load raw data and convert 'date' field data to date data type
```{r}
temp <- tempfile()
# Download file from source if required and write to temp file
#download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp, mode="wb")

# Unzip temp file to working csv for processing
#unzip(temp, "activity.csv")

activityData <- read.csv('activity.csv', header = TRUE, sep = ",",
                  colClasses=c("numeric", "character", "numeric"))

# Cast date field data to date type
activityData$date <- as.Date(activityData$date) 

# Inspect the data
#str(activityData)
#head(activityData)

```



## What is mean total number of steps taken per day?
Remove na values per requirement (For this part of the assignment, you can ignore the missing values in the dataset)
Generate histogram of the total number of steps taken per day

```{r}
cleanActivityData <- na.omit(activityData) 

# sum steps by date
subDailySteps <- rowsum(cleanActivityData$steps, format(cleanActivityData$date, '%Y-%m-%d')) 
dfDailySteps <- data.frame(subDailySteps) 
names(dfDailySteps) <- ("steps")

# Plot histogram
hist(dfDailySteps$steps, 
     main=" ",
     breaks=10,
     xlab="Total Number of Steps Taken Daily")

```

## Output mean and median of daily steps

```{r}
# Generate mean of steps
mean(dfDailySteps$steps)

# Generate median of steps
median(dfDailySteps$steps) 
```



## What is the average daily activity pattern?
```{r}
library(plyr)
# Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
subMeanSteps <- ddply(cleanActivityData,~interval, summarise, mean=mean(steps))

# Plot the average
qplot(x=interval, y=mean, data = subMeanSteps,  geom = "line",
      xlab="5-Minute Interval (24 hr clock)",
      ylab="Steps Count",
      main="Average # Steps Averaged Across All Days"
      )


```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
# Max steps interval
subMeanSteps[which.max(subMeanSteps$mean), ]
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs) (use is.na function)
```{r}
missingValues <- sum(is.na(activityData$steps))

# Output value
missingValues

```

### Strategy for filling in all of the missing values in the dataset. We will replace missing values with the mean of the same interval across all days

```{r}
# fill for an interval. data is the unadulterated data frame and meanSteps (subMeanSteps) is the subsetted dataframe of mean steps
## Create the average number of steps per weekday and interval
avgData <- ddply(cleanActivityData, .(interval, date), summarize, Avg = mean(steps))

## Create dataset with all NAs for substitution
naData<- activityData[is.na(activityData$steps),]
## Merge NA data with average weekday interval for substitution
filledData<-merge(naData, avgData, by=c("interval", "date"))
```

```{r}
newData<- filledData[,c(3,2,1)]
colnames(newData)<- c("steps", "date", "interval")

##Merge the NA averages and non NA data together
mergeData <- rbind(cleanActivityData, newData)




```
Mean
```{r}
##Create sum of steps per date for comparison
sumData <- aggregate(mergeData$steps ~ mergeData$date, FUN=sum, )
colnames(sumData)<- c("Date", "Steps")

## Mean of Steps with NA data taken care of
as.integer(mean(sumData$Steps))
```
Median
```{r}
as.integer(median(sumData$Steps))
```

```{r}
hist(sumData$Steps, breaks=10, xlab="Steps", main = "Total Steps per Day w/ filled data")

```
No difference any consequence for imputing missing values from the original 


## Are there differences in activity patterns between weekdays and weekends?


```{r}
hist(sumData$steps, 
     main=" ",
     breaks=10,
     xlab="Total Number of Steps Taken Daily")


```


